Real-time resolving assembly graph by ONT long reads data

### 2019-11-20 paper.tex
proof-reading

%Display watermark DRAFT
%\watermark{DRAFT}

@@

%Set abstract (see below)
\abstract{
A real-time assembly pipeline for Oxford Nanopore Technology (ONT) data is important for either saving sequencing resources or reducing turnaround time for data analyses. The previous approach from \npscarf{} provided a greedy fast-response, auto-correct streaming algorithm for such task but was relatively prone to mis-assemblies compared to other graph-based methods. 
Here we present \npgraph{}, a real-time hybrid assembly software using the assembly graph instead of the separated pre-assembly contigs. It is able to produce more complete genome assembly by resolving the path finding problem on the assembly graph using long reads as the traversing guide. 
The application on synthetic and real data set of isolate genomes show improved accuracy while still maintaining the required computational cost extremely low. On top of that, we apply \npgraph{} on mock community in the very first application to improve metagenomics assembly in real-time with nanopore data. 
With \npscarf{}, the built-in graphical user interface (GUI) can provide users a comprehensive look-and-feel of the whole assembly process. The tool and source code is available at \url{https://github.com/hsnguyen/assembly}.
%=======
%A real-time assembly pipeline for Oxford Nanopore Technology (ONT) data is important for either better control of sequencing resources or quick turnaround of analyses. The previous approach from \npscarf{} provided a greedy fast-response, auto-correct streaming algorithm for such task but was relatively prone to mis-assemblies. Here we present another real-time hybrid assembler, \npgraph{}, that can work on the assembly graph instead of the pre-assembled contigs. The results show improved accuracy while still minimize the required computational cost. The tool is available at \url{https://github.com/hsnguyen/assembly}.
%>>>>>>> 2d4c481dd93902dfb2dabe46534a62e8a8c2058c
}

%Set keywords
\keywords{hybrid assembly, assembly graph, real-time analysis, nanopore sequencing}

@@

Table~\ref{table:npgraph_compare} shows comparative results running different methods on 5 synthetic data sets, the output of the full run can be found in Supplementary Table~\ref{supp_tab:synthetic_benchmark}.
In the first column of applied methods, beside \unicycler{} and $\mathtt{hybridSPAdes}$, the original scaffolder \npscarf{} was included as well as \npscarfg{} -- its modified version with assembly graph integrated.
On the other hand, \npgraph{} can use 2 different aligners, \bwa{} and \minimap{}, for its bridging phase thus both practices were included in this comparison.

@@

To align the long reads to the assembly graph components, either \bwa{}~\cite{Li2013} or \minimap{}~\cite{Li2016} was invoked in \npgraph{}. 
The former option was inherited from \npscarf{} pipeline with the intact parameters
%$-k11 -W20 -r10 -A1 -B1 -O1 -E1 -L0 -a -Y$
while the latter was used with the recommended settings (-$\mathtt{k}$15 -$\mathtt{w}$5) for the best sensitivity working on MinION data.
Even though, \bwa{} normally reported more hits than \minimap{} but at the same time, was responsible for more false positive alignments.
For instance, regarding the last data set from Table~\ref{table:npgraph_compare}, the assembly of \npgraph{} using \bwa{} was suffered from the ambiguous alignments thus more fragmented than the other counterparts. 
On the other hand, referring to more complicated graphs from \emph{Acinetobacter} A1 and the yeast \emph{Saccharomyces cerevisiae} S288c from Table~\ref{supp_tab:synthetic_benchmark}, the number of mis-assemblies from using \minimap{} were increased due to the lacks of appropriate alignments to support accurate bridging process.
However, under almost circumstances, using either aligners would result in final assemblies with similar qualities.
Furthermore, in terms of running time and resources required, \minimap{} proved to be the best option. 
The total CPU hours had been cut down drastically with the new aligner, making \npgraph{} the fastest hybrid assembler available.
This feature is certainly more favoured to a real-time assembly as well.
Because \minimap{} is expected to replace \bwa{} for long-read sequencing data alignment, it likewise being chosen as the default aligner for \npgraph{} pipeline.

To evaluate assembly errors, we considered the measure of the total mismatches and indels per 100 Kbp from the assembly sequences when mapping to the reference.
As expected from hybrid assembly method embedding Illumina data, the figures were hardly bigger than 100 (equivalent to 0.1\% error rate) in almost every cases.
Furthermore, the indels errors, which mainly caused by TGS data, were found relatively low in the final contigs (Table~\ref{supp_tab:synthetic_benchmark}).
The majority of the differences accounted for the mismatched nucleotides caused by the alternative paths connecting the unique anchors from the backbone of the assembly.
This phenomenon may root from homologous repeats or sequencing errors of the genome.
From all the hybrid assemblers, $\mathtt{hybridSPAdes}$ reported results with highest fidelity. This meant that the performance its decision-rule algorithm $\mathtt{exSPAnder}$~\cite{Prjibelski2014} was the most accurate amongst all path finding methods. As the trade-off, there were fewer connections satisfying its quality threshold, resulting in the fragmented assemblies in cases of \emph{Streptococcus suis} or \emph{Acinetobacter} samples (Table~\ref{table:npgraph_compare} and \ref{supp_tab:synthetic_benchmark}).
\unicycler{}, which employs an algorithm based on semi-global (or glocal) alignments~\cite{Brudno2003glocal} with the consensus long reads, returned the second best reliable and at the same time, closest-to-complete results overall.
\npscarf{}, on the other hand, exploited the long reads for the gap filling thus inherited the high error rates from them.
By integrating the assembly graph for the task, the errors were reduced in general (random sequences, \emph{M.~tuberculosis}, \emph{S.suis} from Table~\ref{table:npgraph_compare}) but not completely since the mis-placed contigs were still not resolved in other circumstances.
\npgraph{} significantly reduced the errors compared to \npscarf{}, however the figures were still higher than the those of the best counterparts.
This implied a more robust decision making system is needed in \npgraph{}'s real-time path finding module for even better output's accuracy.

>>>>

@@

%Set abstract (see below)
\abstract{
A real-time assembly pipeline for Oxford Nanopore Technology (ONT) data is important for either saving sequencing resources or reducing turnaround time for data analyses. The previous approach from \npscarf{} provided a greedy fast-response, auto-correct streaming algorithm for such task but was relatively prone to mis-assemblies compared to other graph-based methods. 
Here we present \npgraph{}, a real-time hybrid assembly software using the assembly graph instead of the separated pre-assembly contigs. It is able to produce more complete genome assembly by resolving the path finding problem on the assembly graph using long reads as the traversing guide. 
The application on synthetic and real data set of isolate genomes show improved accuracy while still maintaining the required computational cost extremely low. On top of that, we apply \npgraph{} on mock community in the very first application to improve metagenomics assembly in real-time with nanopore data. 
With \npscarf{}, the built-in graphical user interface (GUI) can provide users a comprehensive look-and-feel of the whole assembly process. The tool and source code is available at \url{https://github.com/hsnguyen/assembly}.
} 
%Set keywords
\keywords{hybrid assembly, assembly graph, real-time analysis, nanopore sequencing}

@@

Table~\ref{table:npgraph_compare} shows comparative results running different methods on 5 synthetic data sets, report for the whole comparison can be found in Supplementary Table~\ref{supp_tab:synthetic_benchmark}.
In the first column of applied methods, beside \unicycler{} and $\mathtt{hybridSPAdes}$, the original scaffolder \npscarf{} was included as well as \npscarfg{} -- its modified version with assembly graph integrated.
On the other hand, \npgraph{} can use 2 different aligners, \bwa{} and \minimap{}, for its bridging phase thus both practices were included in this comparison.

@@

To align the long reads to the assembly graph components, either \bwa{}~\cite{Li2013} or \minimap{}~\cite{Li2016} was invoked in \npgraph{}. 
The former option was inherited from \npscarf{} pipeline with the intact parameters
%$-k11 -W20 -r10 -A1 -B1 -O1 -E1 -L0 -a -Y$
while the latter was used with the recommended settings (-$\mathtt{k}$15 -$\mathtt{w}$5) for the best sensitivity working on MinION data.
Even though, \bwa{} normally reported more hits than \minimap{} but at the same time, was responsible for more false positive alignments.
For instance, regarding the last data set from Table~\ref{table:npgraph_compare}, the assembly of \npgraph{} using \bwa{} was suffered from the ambiguous alignments thus more fragmented than the other counterparts. 
On the other hand, referring to more complicated graphs from \emph{Acinetobacter} A1 and the yeast \emph{Saccharomyces cerevisiae} S288c from Table~\ref{supp_tab:synthetic_benchmark}, the number of mis-assemblies from using \minimap{} were increased due to the lacks of appropriate alignments to support accurate bridging process.
However, under almost circumstances, using either aligners would result in final assemblies with similar qualities.
Furthermore, in terms of running time and resources required, \minimap{} proved to be the best option. 
The total CPU hours had been cut down drastically with the new aligner, making \npgraph{} the fastest hybrid assembler available.
This feature is certainly more favoured for a real-time assembly pipeline that normally requires quick responses in very short intervals.
Because \minimap{} is expected to replace \bwa{} for long-read sequencing data alignment, it likewise being chosen as the default aligner for \npgraph{} pipeline.

To evaluate assembly errors, we considered the measure of the total mismatches and indels per 100 Kbp from the assembly sequences when mapping to the reference.
As expected from hybrid assembly methods embedding Illumina data, the figures were hardly bigger than 100 (equivalent to 0.1\% error rate) in almost every cases.
Furthermore, the indels errors, which mainly caused by TGS data, were found relatively low in the final contigs (Table~\ref{supp_tab:synthetic_benchmark}).
The majority of the differences accounted for the mismatched nucleotides caused by the alternative paths connecting the unique anchors from the backbone of the assembly.
This phenomenon may root from homologous repeats or sequencing errors of the genome.
From all the hybrid assemblers, $\mathtt{hybridSPAdes}$ reported results with highest fidelity. This meant that the performance its decision-rule algorithm $\mathtt{exSPAnder}$~\cite{Prjibelski2014} was the most accurate amongst all path finding methods. As the trade-off, there were fewer connections satisfying its quality threshold, resulting in the fragmented assemblies in cases of \emph{Streptococcus suis} or \emph{Acinetobacter} samples (Table~\ref{table:npgraph_compare} and \ref{supp_tab:synthetic_benchmark}).
\unicycler{}, which employs an algorithm based on semi-global (or glocal) alignments~\cite{Brudno2003glocal} with the consensus long reads, returned the second best reliable and at the same time, closest-to-complete results overall.
\npscarf{}, on the other hand, employed the long reads for the gap filling thus inherited the high error rates from them.
By integrating the assembly graph for the task, the errors were reduced in general (random sequences, \emph{M.~tuberculosis}, \emph{S.suis} from Table~\ref{table:npgraph_compare}) but not completely since the mis-placed contigs were still not resolved in other circumstances.
\npgraph{} significantly reduced the errors compared to \npscarf{}, however the figures were still higher than the those of the best counterparts.
This implied a more robust decision making system is needed in \npgraph{}'s real-time path finding module for even better output's accuracy.
